---
title: "IS607 - Assignment 12"
author: "Suman K Polavarapu"
date: "Sunday, April 26, 2015"
output: html_document
---

#### Activity : (1) Write the contents of a json file of your choosing into a mongodb database, then (2) Use R to display some of the information from this mongodb database.


__1.Setup - Load data and the required libraries__
Data Source:  http://mysafeinfo.com/api/data?list=presidents&format=json
Import json into MondoDB database "rtest" : C:\\Users\\Suman>mongoimport -d rtest -c presidents < C:\\Users\\Suman\\presidents.json

```{r, warning=FALSE}
#install.packages("rmongodb")
library(rmongodb)
```

__2.Analysis__

```{r, warning=FALSE}
#Connect to my local mongo db server or replset and return an object of class "mongo" for communication over connection.
mgdb <- mongo.create(host="127.0.0.1")

#get collections from my database , which 'rtest'
(mongo.get.database.collections(mgdb, 'rtest'))

#Get Democrat presidents only
bson.query <- mongo.bson.from.JSON('{ "pp" : "Democrat" } ')

cur <- mongo.find(mgdb, "rtest.presidents", bson.query)

democrat.df <- mongo.cursor.to.data.frame(cur)

head(democrat.df)

```

__Now lets try getting all the data from the mongo collection__

```{r}
#Get all presidents
cur.all <- mongo.find(mgdb, "rtest.presidents")

presidents.df <- mongo.cursor.to.data.frame(cur.all)

#Make sure to destroy the connection to mongoDb.
mongo.destroy(mgdb)

names(presidents.df) <- c("id", "name", "party", "time")

head(presidents.df)

```

__Quick visualization - Histogram of US Presidents by party __

```{r}
library(ggplot2)

ggplot(data=presidents.df, aes(x=party, fill = party)) + geom_histogram() + ggtitle("Count of Presidents from each party") +
stat_bin(aes(y=..count.., label=..count..), geom="text", vjust=-.5)

```


